Fréchet Distance

### 2022-12-26 docs/paper.tex
docs: fixed some errors and updated date.

%%% date
\date{December 05, 2022}

@@

% Free Space Diagram
\begin{figure}[ht]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        % \includegraphics[width=\textwidth]{images/frechet_distance/line_segments.pdf}
        \input{images/frechet_distance/line_segments.tex}
      }
      \caption{Line segments $P$ and $Q$}
      \label{fig_free_space_lines_example}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        % \includegraphics[width=\textwidth]{images/frechet_distance/free_space_single_cell.pdf}
        \input{images/frechet_distance/free_space_single_cell.tex}
      }
      \caption{Free space diagram}
      \label{fig_free_space_diagram_example}
  \end{subfigure}
  \caption[Free space diagram example]{
    The free space diagram \autoref{fig_free_space_diagram_example} 
    for line segments $P$ and $Q$ and distance $\varepsilon$
    from \autoref{fig_free_space_lines_example} using euclidean distance.}
  \label{fig_free_space}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        % \includegraphics[width=\textwidth]{images/frechet_distance/polygonal_curves.pdf}
        \input{images/frechet_distance/polygonal_curves.tex}
      }
      \caption{Polygonal curves $P$ and $Q$}
      \label{fig_complex_free_space_lines_example}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        \includegraphics[width=\textwidth]{images/frechet_distance/free_space_polygonal_curves.pdf}
        % \input{images/frechet_distance/free_space_polygonal_curves.tex}
      }
      \caption{Free space diagram}
      \label{fig_complex_free_space_diagram_example}
  \end{subfigure}
  \caption[Complex free space diagram example]{
    The free space diagram \autoref{fig_complex_free_space_diagram_example} 
    for polygonal curves $P$ and $Q$ from \autoref{fig_complex_free_space_lines_example} 
    and distance $\varepsilon = 0.73$ using euclidean distance. (Figure inspired by \cite{buchin_four_2017})}
  \label{fig_complex_free_space}
\end{figure}

@@

The second algorithm works by directly searching for a path from $s$ to $t$ in $G$,
without any prior removal of edges from $E$.
This can be done using Prim's minimum spanning tree algorithm \cite{prim_shortest_1957} starting at $s$ 
and stopping as soon as $t$ is included in the spanning tree. 
The value of $\varepsilon$ is the value of the last added edge.
Similarly, one can use Dijkstra's algorithm \cite{dijkstra_note_1959}, 
optimizing for the minimum maximum edge value.
Both approaches result in a time complexity of $\mathcal{O}(pq \log (pq))$. \cite{alt_computing_1995}
In this case of a planar grid based graph with non-negative edge values 
an even faster algorithm from Henzinger et al. \cite{henzinger_faster_1997} to solve the single source shortest paths 
can be applied. This gives a time complexity of $\mathcal{O}(pq)$ to compute the weak Fréchet distance \cite{van_leusden_novel_2013}.

\subsection{Algorithm for the Discrete Fréchet Distance}
\label{sec_discrete_algorithm}
The computation of the discrete Fréchet distance, as defined in \autoref{def_discrete_frechet_distance}, is very easy and straight forward 
using a dynamic programming approach as originally suggested by Eiter et al. \cite{eiter_computing_1994}.
Recall that the polygonal curves $P = (u_1, \dots, u_p)$ and $Q = (v_1, \dots, v_p)$ are defined as sequences of points.
The algorithm works by recursively computing the minimal required $\varepsilon$ for partial curves 
$P' = (u_1, \dots, u_i)$ and $Q' = (v_1, \dots, v_j)$
using $c(i,j)$ as defined in \autoref{def_recursive_forumla_discrete_frechet_distance}.
Starting from $c(1,1)$, which is the distance between the starting points $u_1$ and $v_1$,
all other values of $c(i,j)$ are computed iteratively in a bottom up manner 
and memoization on the previously computed values of $c(i,j)$.
The resulting algorithm has runtime of $\mathcal{O}(pq)$
because the constant time function $c(i,j)$ is called $pq$ times \cite{eiter_computing_1994}.

\begin{mydef}
  \label{def_recursive_forumla_discrete_frechet_distance}
  The discrete Fréchet distance of polygonal curves $P' = (u_1, \dots, u_i)$ and $Q' = (v_1, \dots, v_j)$
  can be computed by
  \begin{align*}
    c(i,j) &= \begin{cases}
      dist(u_1, v_1), &\text{if } i=1 \text{and } j=1 \\
      \max \{c(i-1,1), dist(u_i, v_1)\}, &\text{if } i>1 \text{and } j=1 \\
      \max \{c(1,j-1), dist(u_1, v_j)\}, &\text{if } i=1 \text{and } j>1 \\
      \max \{c(i-1,j), c(i,j-1), c(i-1,j-1), dist(u_i, v_j)\}, &\text{if } i>1 \text{and } j>1
    \end{cases} \\
    &= \delta_{dF}(P',Q')
  \end{align*}
\end{mydef}

>>>>

%%% date
\date{December 26, 2022}

@@

% Free Space Diagram
\begin{figure}[ht]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        % \includegraphics[width=\textwidth]{images/frechet_distance/line_segments.pdf}
        \input{images/frechet_distance/line_segments.tex}
      }
      \caption{Line segments $P$ and $Q$.}
      \label{fig_free_space_lines_example}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        % \includegraphics[width=\textwidth]{images/frechet_distance/free_space_single_cell.pdf}
        \input{images/frechet_distance/free_space_single_cell.tex}
      }
      \caption{Free space diagram.}
      \label{fig_free_space_diagram_example}
  \end{subfigure}
  \caption[Free space diagram example]{
    The free space diagram \autoref{fig_free_space_diagram_example} 
    for line segments $P$ and $Q$ and distance $\varepsilon = 0.73$
    from \autoref{fig_free_space_lines_example} using euclidean distance.}
  \label{fig_free_space}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        % \includegraphics[width=\textwidth]{images/frechet_distance/polygonal_curves.pdf}
        \input{images/frechet_distance/polygonal_curves.tex}
      }
      \caption{Polygonal curves $P$ and $Q$.}
      \label{fig_complex_free_space_lines_example}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\textwidth}
      \resizebox{\textwidth}{!}{
        \includegraphics[width=\textwidth]{images/frechet_distance/free_space_polygonal_curves.pdf}
        % \input{images/frechet_distance/free_space_polygonal_curves.tex}
      }
      \caption{Free space diagram for a non-minimal value of $\varepsilon$.}
      \label{fig_complex_free_space_diagram_example}
  \end{subfigure}
  \caption[Complex free space diagram example]{
    The free space diagram \autoref{fig_complex_free_space_diagram_example} 
    for polygonal curves $P$ and $Q$ from \autoref{fig_complex_free_space_lines_example} 
    and distance $\varepsilon = 0.73$ using euclidean distance. (Figure inspired by \cite{buchin_four_2017})}
  \label{fig_complex_free_space}
\end{figure}

@@

The second algorithm works by directly searching for a path from $s$ to $t$ in $G$,
without any prior removal of edges from $E$.
This can be done using Prim's minimum spanning tree algorithm \cite{prim_shortest_1957} starting at $s$ 
and stopping as soon as $t$ is included in the spanning tree. 
The value of $\varepsilon$ is the maximum value of all edged added to the spanning tree.
Similarly, one can use Dijkstra's algorithm \cite{dijkstra_note_1959}, 
optimizing for the minimum maximum edge value.
Both approaches result in a time complexity of $\mathcal{O}(pq \log (pq))$. \cite{alt_computing_1995}
In this case of a planar grid based graph with non-negative edge values 
an even faster algorithm from Henzinger et al. \cite{henzinger_faster_1997} to solve the single source shortest paths problem 
can be applied. This gives a time complexity of $\mathcal{O}(pq)$ to compute the weak Fréchet distance \cite{van_leusden_novel_2013}.

\subsection{Algorithm for the Discrete Fréchet Distance}
\label{sec_discrete_algorithm}
The computation of the discrete Fréchet distance, as defined in \autoref{def_discrete_frechet_distance}, is very easy and straight forward 
using a dynamic programming approach as originally suggested by Eiter et al. \cite{eiter_computing_1994}.
Recall that the polygonal curves $P = (u_1, \dots, u_p)$ and $Q = (v_1, \dots, v_p)$ are defined as sequences of points.
The algorithm works by recursively computing the minimal required $\varepsilon$ for partial curves 
$P' = (u_1, \dots, u_i)$ and $Q' = (v_1, \dots, v_j)$
using $c(i,j)$ as defined in \autoref{def_recursive_forumla_discrete_frechet_distance}.
Starting from $c(1,1)$, which is the distance between the starting points $u_1$ and $v_1$,
all other values of $c(i,j)$ are computed iteratively in a bottom up manner 
with memoization on the previously computed values of $c(i,j)$.
The resulting algorithm has runtime of $\mathcal{O}(pq)$
because the constant time function $c(i,j)$ is called $pq$ times \cite{eiter_computing_1994}.
Additionally, the algorithm can be implemented to require $\mathcal{O}(min(p,q))$ space 
because it is sufficient to store the previous column or row of $c(i,j)$ values 
for memoization.

\begin{mydef}
  \label{def_recursive_forumla_discrete_frechet_distance}
  The discrete Fréchet distance of the two (partial) polygonal curves $P' = (u_1, \dots, u_i)$ and $Q' = (v_1, \dots, v_j)$
  can be computed by
  \begin{align*}
    c(i,j) &= \begin{cases}
      dist(u_1, v_1), &\text{if } i=1 \text{and } j=1 \\
      \max \{c(i-1,1), dist(u_i, v_1)\}, &\text{if } i>1 \text{and } j=1 \\
      \max \{c(1,j-1), dist(u_1, v_j)\}, &\text{if } i=1 \text{and } j>1 \\
      \max \{c(i-1,j), c(i,j-1), c(i-1,j-1), dist(u_i, v_j)\}, &\text{if } i>1 \text{and } j>1
    \end{cases} \\
    &= \delta_{dF}(P',Q')
  \end{align*}
\end{mydef}
